Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-11-18 paper/main.tex
rewrite application 1
- adjust to new simulation
- try to explain more clearly what is being done

\textbf{@Christian is this more clear now?}

We model the force exerted by the fluid flow onto the individual vertices by

\begin{equation}
    \vec{F}_i = C |\vec{x}_i - \vec{x}_{i+1}|\sin(\alpha_i)\vec{e}_x
\end{equation}

which acts similarly to Stoke's Law \textbf{TODO Citation} onto the normal component of each
rod-segment where $\alpha_i$ is the angle of the rod edge between $\vec{x}_i$ and $\vec{x}_{i+1}$
and the x-axis.
We extract the position of the rod from the provided image and use it as initial value.
Our simulation is split into two phases, one with an external force that initially bends the rod,
and a second one during which the force is disabled and the cell returns to its original shape.
We compare the simulated positions with the extracted ones and calculate the euclidean distance
between them, which we use as the cost function to estimate the parameters of this model.

Since this model is void of any cell-cell interactions or cell division, we are left with 5 parameters:
Magnitude of the drag force $C$, growth rate $\alpha$, rigidity of the rod $\eta$, spring tension
$\gamma$ and damping constant $\lambda$ (see table~\ref{table:list-of-parameters}).
Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (orange) and final
(blue) position of the rod using the optimized parameters.
While the parameters for the dragging force, growth rate and damping can be identified, the same
can not be said for the rod rigidity and spring tension, which both are not identifiable.
However this behaviour is expected since, the shape of the rod is almost exclusively straight, thus
adding no further information except that a lower threshold of rigidity and spring tension should be
present.

Note that each point represented in the subfigures~\ref{fig:amir-bending-simulation} (C-G) was
optimized again, meaning that higher values of rigidity and spring tension can also lead to higher
values of drag forces.

>>>>

We model the force exerted by the fluid flow onto the individual vertices by
\begin{equation}
    \vec{F}_{i,\text{external}} = C |\vec{x}_i - \vec{x}_{i+1}|(y-y_b)\sin(\beta_i)\vec{e}_x
\end{equation}
which is proportional to the length of the rod segment projected along the y-axis.
It acts similarly to Stoke's Law~\cite{Stokes1856} along the axis of the fluid flow (x-axis).
Here $\beta_i$ is the angle of the rod edge between $\vec{x}_{i+1}-\vec{x}_i$ and the x-axis.
Amir et. al~\cite{Amir2014_2} calculated the flow profile for this particular setup.
Since the bacterium does not reach the peak of the velocity profile, we can estimate the slope via a
linear dependence on the pertruding segment along the y-axis~\cite{Wolschin2021}.
In particular, this enforces the correct boundary values the flow profile and the resulting force
$\vec{F}_{i,\text{external}}|_{y=0}=0$.

We extract the positions of the rod from the data images (see
Figure~\ref{fig:amir-bending-simulation} (A)) and compare the simulated positions with the extracted
ones using the direct comparison algorithm (see Subsection~\ref{subsec:direct-comparison}).
Our simulation is split into two phases; one with an external force that initially bends the rod,
and a second one during which the force is disabled and the cell returns to its original shape.
We assume that these time intervals are identically long since the data from the movies does not
allow for a finer resolution of the dynamics.

Since this model is void of any cell-cell interactions or cell division, the dynamics of the cell
can be described by a combination of growth and physical mechanics (see
subsection~\ref{subsection:mechanical-model-mechanics}).
We are thus left with 5 parameters:
Magnitude of the drag force $C$, growth rate $\alpha$, rigidity of the rod $\eta$, spring tension
$\gamma$ and damping constant $\lambda$ (see table~\ref{table:list-of-parameters}).
Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (orange) and final
(blue) position of the rod using the optimized parameters.
While the parameters for the dragging force, growth rate and damping can be identified, the same
can not be said for the rod rigidity and spring tension, which both are not identifiable.
However this behaviour is expected since, the shape of the rod is almost exclusively straight, thus
adding no further information except that a lower threshold of rigidity and spring tension should be
present.
